---
title: "HW1_jaeyounglee"
author: "Jaeyoung Lee"
date: \today
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1
I have practiced the visualization and programming basics in Primers

# Problem 2 
## Part A

What I hope to get out of this class as follows, 
\begin{enumerate}
\item How to be reproducible to collaborate with others and even myself 
\item How to write effective codes
\item How to make nice reports using R Markdown (Be a nice statistician) 
\item Getting used to R Markdown and GitHub 
\end{enumerate}
I feel lucky to take this class.


## Part B
### Binomial, Gamma, Beta pdfs from Casella & Berger


 
\begin{eqnarray}
P(X = x | n, p) &=& {n \choose x} p^x (1-p)^{1-x}. \\
P(X|\alpha, \beta) &=& \frac{1}{\gamma(\alpha)\beta^\alpha} x^{\alpha -1} e^{-\frac{x}{\beta}}. \\
P(X|\alpha, \beta) &=& \frac{1}{B(\alpha, \beta)} x^{\alpha-1} (1-x)^{\beta-1}.
\end{eqnarray} 




# Problem 3

With Rule 1, it is useful to make comments for every steps. There is no need to use rule 2 for problem 4, but in the future, it is essential not to use Excel or other programs for data manipulation. From Rule 3, it is essential to write the version of R console whenever we make different version of R code to avoid errors. Using GitHub, we can follow the rule 4. According to rule 5, it is better to record intermediate results, so that we can fix problems easily. To control randomness and show the same result, setting a random seed is important by rule 6. For readability, following rule 7 is helpful. R Markdown is useful to follow rule 8. ReadMe file is a good example for rule 9. For rule 10, GitHub is again a great tool. 




# Problem 4

```{r problem4}

### Problem 4 : A Scatter Plot and A Histogram ###

# R version 4.0.2
library(help = 'datasets')  # To get a list of the datasets
summary(trees)              # Summary of trees data

### A basic scatter plot from 'trees' data using ggplot2
library(ggplot2)            # To use ggplot function
ggplot(data = trees) + 
  geom_point(mapping = aes(x = Height, y = Volume, color = Girth))

### Histogram for 'trees' data
hist(trees$Volume, main ='Histogram of Volume of Trees', xlab = 'Volume')


```

# Problem 5

Push to GitHub

\pagebreak


# Appendix : R codes
```{r Appendix, ref.label=c("setup","problem4"), echo=TRUE, eval=FALSE, tidy=FALSE, include=T}
  
```
