---
title: "HW1_jaeyounglee"
author: "Jaeyoung Lee"
date: \today
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1
I have practiced the visualization and programming basics in Primers

# Problem 2 
## Part A

What I hope to get out of this class as follows, 
\begin{enumerate}
\item How to be reproducible to collaborate with others and even myself 
\item How to write effective codes
\item How to make nice reports using R Markdown (Be a nice statistician) 
\item Getting used to R Markdown and GitHub 
\end{enumerate}
I feel lucky to take this class.


## Part B
### Binomial, Gamma, Beta pdfs from Casella & Berger

\begin{eqnarray}
P(X = x | n, p) &=& {n \choose x} p^x (1-p)^{1-x}. \\
P(X|\alpha, \beta) &=& \frac{1}{\gamma(\alpha)\beta^\alpha} x^{\alpha -1} e^{-\frac{x}{\beta}}. \\
P(X|\alpha, \beta) &=& \frac{1}{B(\alpha, \beta)} x^{\alpha-1} (1-x)^{\beta-1}.
\end{eqnarray} 

\pagebreak


# Problem 3

\begin{enumerate}
  \item Rule 1. For Every Result, Keep Track of How It Was Produced
    \begin{itemize}
      \item Recording every detail is important for reproducibility. 
      \item Details such as the name and version of the program, parameters and inputs are critical.
    \end{itemize}
    
  \item Rule 2. Avoid Manual Data Manipulation Steps
    \begin{itemize}
      \item Rely on codes, not on manual procedures such as Excel to reproduce.
    \end{itemize}
    
  \item Rule 3. Archive the Exact Versions of All External Programs Used
    \begin{itemize}
      \item A newer version of a program may not run even without any change of inputs.
    \end{itemize}
    
  \item Rule 4. Version Control All Custom Scripts
    \begin{itemize}
      \item Only one exact script may be able to produce that exact output.
      \item Sometimes, backtracking to a code state is needed.
      \item Use GitHub.
    \end{itemize}
    
  \item Rule 5. Record All Intermediate Results, When Possible in Standardized Formats
    \begin{itemize}
      \item Looking through intermediate results can uncover discrepancies toward what is assumed, and can find bugs or faulty interpretations. 
      \item It reveals consequences of alternative programs and parameter choices at each steps. 
      \item It allows parts of the process to be rerun. 
      \item One can track the steps where the problems appears. 
      \item Without full operation, one can examine the full process. 
    \end{itemize}
    
  \item Rule 6. For Analyses That Include Randomness, Note Underlying Random Seeds
    \begin{itemize}
      \item Providing the same random seed allows results to be reproduced exactly in future runs.
    \end{itemize}
    
  \item Rule 7. Always Store Raw Data behind Plots
    \begin{itemize}
      \item It allows raw data for a given figure to be easily retrieved.
      \item one can easily modify the plotting procedure, without redoing the whole analysis.
    \end{itemize}
    
  \item Rule 8. Generate Hierarchical Analysis Output, Allowing Layers of Increasing Detail to Be Inspected
    \begin{itemize}
      \item Hypertext is the best example.
      \item By simple clicks, we can easily view the full data underlying the summary of results with links.
    \end{itemize}
    
  \item Rule 9. Connect Textual Statements to Underlying Results
    \begin{itemize}
      \item Connect results to the statements that are initially formulated such as notes or emails.
      \item It is important to provide details along with your textual interpretations to the results to be tracked down in the future.
    \end{itemize}
    
  \item Rule 10. Provide Public Access to Scripts, Runs, and Results
    \begin{itemize}
      \item All inputs, scripts, versions, parameters, and intermediate results should be provided publicly and easily accessible.
      \item GitHub is a good tool.
    \end{itemize}
\end{enumerate}


\pagebreak


# Problem 4

```{r problem4}

### Problem 4 : A Scatter Plot and A Histogram ###

# R version 4.0.2
library(help = 'datasets')  # To get a list of the datasets
summary(trees)              # Summary of trees data

### A basic scatter plot from 'trees' data using ggplot2
library(ggplot2)            # To use ggplot function
ggplot(data = trees) + 
  geom_point(mapping = aes(x = Height, y = Volume, color = Girth))

### Histogram for 'trees' data
hist(trees$Volume, main ='Histogram of Volume of Trees', xlab = 'Volume')


```

# Problem 5

Push to GitHub

\pagebreak


# Appendix : R codes
```{r Appendix, ref.label=c("setup","problem4"), echo=TRUE, eval=FALSE, tidy=FALSE, include=T}
  
```
