---
title: "HW2_jaeyounglee"
author: "Jaeyoung Lee"
date: \today
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(ggplot2)
library(ggpubr)
```
## Problem 3

First of all, it is to handle a mistake. Also, one can handle various versions of a code and see the history of a code. Furthermore, using version control, it is easy to collaborate with others. This is because one can share a code and work on the cloud such as GitHub.



## Problem 4

a. Sensory data from five operators.
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat> 
```{r sensory_import}
######## Sensory data ########
# Getting "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
url_sensory <- "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensory_rawdata <- fread(url_sensory, fill = TRUE, skip = 2, data.table = FALSE)
saveRDS(sensory_rawdata, 'sensory_rawdata.RDS')
sensory_rawdata <- readRDS('sensory_rawdata.RDS')

```

There are missing values in the raw data and the categories "Items" are in the data like oberservations. We need to remove missing values and extract the 'Item' numbers from the data.

```{r sensory_base}
# Using base R function only
# Convert data.frame to matrix and transpose the raw data
matrix_sensory <- t(as.matrix(sensory_rawdata)) 

# Find where the missing values are
na <- which(is.na(matrix_sensory==TRUE))        

# The indexes where Item numbers are in the data
x <- 1
item <- x
for (i in 1:9){
  x <- x+18
  item <- c(item, x)
}

# Remove missing values and 'Item's from the data
sensory_data <- t(matrix(matrix_sensory[-c(na,item)], byrow = T, nrow = 10))  
sensory_data <- data.table(sensory_data)                                      
colnames(sensory_data) <- paste('Item', 1:10) # Assign column names
Opr <- rep(paste('Opr', 1:5), 3)              # Operator names
sensory_data <- cbind(Opr,sensory_data)       # Bind Operator names and the data

# Re-order the rows by names of operators
sensory_data <- sensory_data[order(sensory_data$Opr)]                         
sensory_data_base <- sensory_data

# Final tidy data with base R functions
head(sensory_data_base)

```

Above is the converted tidy data frames using the base R functions only. A summary of the data is as follows:

```{r summary_sensory_base, echo = FALSE}
knitr::kable(summary(sensory_data_base))
```

\pagebreak


Now, handle the same data with `tidyverse` package.

```{r sensory_tidyverse}
# Sensory data with tidyverse package
# Making matrix which is the same with base R function but using pipes.
matrix_sensory <- sensory_rawdata %>% as.matrix() %>% t() 
na <- which(is.na(matrix_sensory==TRUE))     # Find missing values

# The indexes where Item numbers are in the data
x <- 1
item <- x
for (i in 1:9){
  x <- x+18
  item <- c(item, x)
}

# Remove missing values and Item numbers from the data
sensory_data <- matrix_sensory[-c(na,item)] %>% matrix(byrow = T, nrow = 10) %>% t()
sensory_data <- data.table(sensory_data)
Opr <- rep(paste('Opr', 1:5), 3)
sensory_data <- bind_cols(Opr,sensory_data) # bind operators and data
colnames(sensory_data) <- c('Opr',paste('Item', 1:10))

# Re-order the rows by names of operators
sensory_data <- sensory_data[order(sensory_data$Opr)] 
sensory_data_tidyverse <- sensory_data

# Final tidy data with tidyverse
sensory_data_tidyverse %>% head()

```

\pagebreak

The result by tidyverse is the same with the base R function. The summary of the data converted by tidyverse is as follows.

```{r summary_sensory_tidayverse, echo = FALSE}
knitr::kable(summary(sensory_data_tidyverse))
```

Here are some box plots of the sensory data. Gray plots are based on base R functions and green ones are from tidyverse.

```{r sensory_plots1, echo = FALSE}
# box plots for item 1 to 4
base_item1 <- ggplot(sensory_data_base, aes(x=Opr, y= sensory_data_base[[2]])) + 
  geom_boxplot(fill = 'gray') + 
  labs(title="Box plot of Item 1",x="Operators", y = "Item 1")

base_item2 <- ggplot(sensory_data_base, aes(x=Opr, y= sensory_data_base[[3]])) + 
  geom_boxplot(fill = 'gray') + 
  labs(title="Box plot of Item 2",x="Operators", y = "Item 2")

base_item3 <- ggplot(sensory_data_base, aes(x=Opr, y= sensory_data_base[[4]])) + 
  geom_boxplot(fill = 'gray') + 
  labs(title="Box plot of Item 3",x="Operators", y = "Item 3")

base_item4 <- ggplot(sensory_data_base, aes(x=Opr, y= sensory_data_base[[5]])) + 
  geom_boxplot(fill = 'gray') + 
  labs(title="Box plot of Item 4",x="Operators", y = "Item 4")

ggarrange(base_item1, base_item2, base_item3, base_item4, ncol = 2, nrow = 2)

```


```{r sensory_plots2, echo = FALSE}
tidy_item5 <- ggplot(sensory_data_tidyverse, aes(x=Opr, y= sensory_data_base[[6]])) + 
  geom_boxplot(fill = 'chartreuse3') + 
  labs(title="Box plot of Item 5",x="Operators", y = "Item 5")

tidy_item6 <- ggplot(sensory_data_tidyverse, aes(x=Opr, y= sensory_data_base[[7]])) + 
  geom_boxplot(fill = 'chartreuse3') + 
  labs(title="Box plot of Item 6",x="Operators", y = "Item 6")

tidy_item7 <- ggplot(sensory_data_tidyverse, aes(x=Opr, y= sensory_data_base[[8]])) + 
  geom_boxplot(fill = 'chartreuse3') + 
  labs(title="Box plot of Item 7",x="Operators", y = "Item 7")

tidy_item8 <- ggplot(sensory_data_tidyverse, aes(x=Opr, y= sensory_data_base[[9]])) + 
  geom_boxplot(fill = 'chartreuse3') + 
  labs(title="Box plot of Item 8",x="Operators", y = "Item 8")

ggarrange(tidy_item5, tidy_item6, tidy_item7, tidy_item8, ncol = 2, nrow = 2)
```

```{r sensory_plots3, echo = FALSE}
base_item9 <- ggplot(sensory_data_base, aes(x=Opr, y= sensory_data_base[[10]])) + 
  geom_boxplot(fill = 'gray') + 
  labs(title="Box plot of Item 9",x="Operators", y = "Item 9")

tidy_item10 <- ggplot(sensory_data_tidyverse, aes(x=Opr, y= sensory_data_base[[11]])) + 
  geom_boxplot(fill = 'chartreuse3') +
  labs(title="Box plot of Item 10",x="Operators", y = "Item 10")

ggarrange(base_item9, tidy_item10, ncol = 2, nrow = 1)
```


b. Gold Medal performance for Olympic Men's Long Jump, year is coded as 1900=0.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat>  
```{r medal_import}
######## Long Jump data ########
# Getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
url_medal <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
medal_rawdata <- fread(url_medal)
saveRDS(medal_rawdata, 'medal_rawdata.RDS')
medal_rawdata <- readRDS('medal_rawdata.RDS')

```

The raw data has missing values and wide type data. It is better to reshape the data.
Also, we need two vectors : 'Year' and 'Long Jump'.

```{r medal_base}
# Using base R function only
# Year is coded as 1900 = 0
# Extract Year and Long Jump vectors
year <- c(medal_rawdata[[1]], medal_rawdata[[3]], 
          medal_rawdata[[5]], medal_rawdata[[7]]) + 1900
longjump <- c(medal_rawdata[[2]], medal_rawdata[[4]], 
              medal_rawdata[[6]], medal_rawdata[[8]])

# Bind the vectors as a data table and rename the categories
medal_data <- data.table(year[1:(length(year)-2)], 
                         longjump[1:(length(longjump)-2)])
colnames(medal_data) <- c('Year', 'LongJump')
medal_data_base <- medal_data

# Final tidy data with base R functions
head(medal_data_base)

```

Above is the converted tidy data frames using the base R functions. A summary of the data is as follows:

```{r summary_medal_base, echo = FALSE}
knitr::kable(summary(medal_data_base))
```

\pagebreak

Now, handle the same data with `tidyverse` package.

```{r medal_tidyverse}
# Using tidyverse package
# Year is coded as 1900 = 0
medal_data <- medal_rawdata[,1:8] # remove missing values only columns

# Extracting 'Year' columns and 'Long Jump' columns and remove missing values
colnames(medal_data) <- paste(rep(c('Year', 'Jump'),4), rep(1:4,each = 2))
year <- medal_data[,c(1,3,5,7)] %>% 
  gather(key = 'name1', value = 'Year', 1,2,3,4) %>% filter(Year != na)
year[,2] <- year[,2] + 1900
jump <- medal_data[,c(2,4,6,8)] %>% 
  gather(key = 'name2', value = 'LongJump', 1,2,3,4) %>% filter(LongJump != na)

# Bind the vectors as a data table and rename the categories
medal_data <- bind_cols(year[,2], jump[,2])
colnames(medal_data) <- c('Year', 'LongJump')
medal_data_tidyverse <- medal_data

# Final tidy data with tidyverse
medal_data_tidyverse %>% head()


```

The result by tidyverse is the same with the base R function. The summary of the data converted by tidyverse is as follows.

```{r summary_medal_tidyverse, echo = FALSE}
knitr::kable(summary(medal_data_tidyverse))
```

\pagebreak

As informative plots, line plots are used.

```{r medal_plots, echo = FALSE}
# Plot of Long Jump by Years
medal_base_plot <- ggplot(data = medal_data_base, aes(x = Year, y= LongJump)) + 
  geom_line(col = 'brown') + 
  labs(title = 'Plot of Long Jump by Years with Base R')

medal_tidyverse_plot <- ggplot(data = medal_data_tidyverse, aes(x = Year, y= LongJump)) + 
  geom_line(col = 'cornflowerblue') + 
  labs(title = 'Plot of Long Jump by Years with tidyverse')

ggarrange(medal_base_plot, medal_tidyverse_plot, ncol = 1, nrow = 2)

```


\pagebreak


c. Brain weight (g) and body weight (kg) for 62 species.    
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat>  
```{r brain_import}
######## Brain weight data ########
# Getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
url_brain <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
brain_rawdata <- fread(url_brain)
saveRDS(brain_rawdata, 'brain_rawdata.RDS')
brain_rawdata <- readRDS('brain_rawdata.RDS')
```

The data needs two columns which are 'Body Wt' and 'Brain Wt'.

```{r brain_base}
# Using base R function only
# The method is the same with the data from part (b)
# Extract Body Wt and Brain Wt vectors
bodywt <- c(brain_rawdata[[1]], brain_rawdata[[3]], brain_rawdata[[5]])
brainwt <- c(brain_rawdata[[2]], brain_rawdata[[4]], brain_rawdata[[6]])

# Remove missing values
brain_data <- data.table(bodywt[-length(bodywt)], brainwt[-length(brainwt)])
colnames(brain_data) <- c('Body_Wt', 'Brain_Wt')
brain_data_base <- brain_data

# Final tidy data with base R functions
head(brain_data_base)

```

Above is the converted tidy data frames using the base R functions. A summary of the data is as follows:

```{r summary_brain_base, echo = FALSE}
knitr::kable(summary(brain_data_base))
```

\pagebreak

Now, handle the same data with `tidyverse` package.

```{r brain_tidyverse}
# Tidy data with tidyverse
# Remove vectors which have only missing values
brain_data <- brain_rawdata[,1:6]

# Extracting 'Year' columns and 'Long Jump' columns and remove missing values
colnames(brain_data) <- paste(rep(c('bw', 'brw'),3), rep(1:3,each = 2))
bw <- brain_data[,c(1,3,5)] %>% gather(key = 'name1', value = 'BW', 1,2,3) %>% 
  filter(BW != na)
brw <- brain_data[,c(2,4,6)] %>% gather(key = 'name2', value = 'BRW', 1,2,3) %>% 
  filter(BRW != na)

# Bind the vectors as a data table and rename the categories
brain_data <- bind_cols(bw[,2], brw[,2])
colnames(brain_data) <- c('Body_Wt', 'Brain_Wt')
brain_data_tidyverse <- brain_data

# Final tidy data with tidyverse
brain_data_tidyverse %>% head()

```

The result by tidyverse is the same with the base R function. The summary of the data converted by tidyverse is as follows.

```{r summary_brain_tidyverse, echo = FALSE}
knitr::kable(summary(brain_data_tidyverse))
```

\pagebreak

As informative plots, scatter plots are used.

```{r brain_plots, echo = FALSE}
# Scatter plots of "Body Wt" and "Brain Wt"
brain_base_plot <- 
  ggplot(data = brain_data_base, aes(x = Body_Wt, y= Brain_Wt)) +
  geom_point(col = 'brown', size = 3, shape = 19) +
  labs(title = 'Using base')

brain_tidyverse_plot <-   
  ggplot(data = brain_data_tidyverse, aes(x = Body_Wt, y= Brain_Wt)) +
  geom_point(col = 'cornflowerblue', size = 3, shape = 19) +
  labs(title = 'Using tidyverse')

# Scatter plots without outliers
brain_base_plot_wo_outlier <-   
  ggplot(data = brain_data_base[-c(19,33),], aes(x = Body_Wt, y= Brain_Wt)) +
  geom_point(col = 'brown', size = 3, shape = 19) +
  labs(title = 'Base without outlier')

brain_tidyverse_plot_wo_outlier <-   
  ggplot(data = brain_data_tidyverse[-c(19,33),], aes(x = Body_Wt, y= Brain_Wt)) +
  geom_point(col = 'cornflowerblue', size = 3, shape = 19) +
  labs(title = 'Tidyverse without outlier')



ggarrange(brain_base_plot, brain_tidyverse_plot, 
          brain_base_plot_wo_outlier, 
          brain_tidyverse_plot_wo_outlier,
          ncol = 2, nrow = 2)

```



\pagebreak




d. Triplicate measurements of tomato yield for two varieties of tomatos at three planting densities.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat>  
```{r tomato_import} 
######## Tomato data ########
# Getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
url_tomato <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
tomato_rawdata <- fread(url_tomato, skip = 1)
saveRDS(tomato_rawdata, 'tomato_rawdata.RDS')
tomato_rawdata <- readRDS('tomato_rawdata.RDS')

```

The values are grouped in the cells of the data above. Therefore, we should split the cells into single values. 

```{r tomato_base}
# Using base R function only
# Need to split the values
cells <- strsplit(unlist(tomato_rawdata), split = ',', fixed = T) # split the data
values <- as.numeric(unlist(c(cells[3:8]))) # numerical data

# Combine the split values into data frame
tomato_matrix <- matrix(values, byrow = T, ncol = 3)
tomato_matrix <- t(cbind(tomato_matrix[1:2,], tomato_matrix[3:4,], tomato_matrix[5:6,]))

# Bind the data with the densities (categories)
tomato_data <- data.frame(tomato_matrix, as.character(rep(c(10000,20000,30000), each=3)))
colnames(tomato_data) <- c('IFE1', 'PusaEarlyDwarf','Density')
tomato_data_base <- tomato_data

# Final tidy data with base R functions
tomato_data_base

```
Above is the converted tidy data frames using the base R functions. A summary of the data is as follows:

```{r summary_tomato_base, echo = FALSE}
knitr::kable(summary(tomato_data_base))
```

Now, handle the same data with `tidyverse` package.

```{r tomato_tidyverse}
# Using tidyverse package
# Need to split the values
tomato_data <- tomato_rawdata[,-1] %>% 
  separate(col = '10000', into = c("1","2","3"), sep = ",", convert = T) %>%
  separate(col = '20000', into = c("4","5","6"), sep = ",", convert = T) %>%
  separate(col = '30000', into = c("7","8","9"), sep = ",", convert = T) %>% 
  as.matrix() %>% t()
dens <- rep(c(10000,20000,30000), each = 3) %>% as.character() # Densities

# Bind the data with the densities vector
tomato_data <- tomato_data %>% data.frame(dens)
colnames(tomato_data) <- c('IFE1', 'PusaEarlyDwarf','Density')
tomato_data_tidyverse <- tomato_data

# Final tidy data with tidyverse
tomato_data_tidyverse

```


```{r summary_tomato_tidyverse, echo = FALSE}
knitr::kable(summary(tomato_data_tidyverse))
```

\pagebreak

Here are the histograms of above data based on both base R and tidyverse.

```{r tomato_plots, echo = FALSE}
# Histograms of tomato data

hist_ife_base = ggplot(tomato_data_base, aes(x=IFE1)) + 
  geom_histogram(binwidth = 1, fill=I("cornflowerblue"), col = I("black")) +
  labs(title = 'Using Base R')

hist_pusa_base = ggplot(tomato_data_base, aes(x=PusaEarlyDwarf)) + 
  geom_histogram(binwidth = 2, fill=I("gold"), col = I("black")) +
  labs(title = 'Using Base R')

hist_ife_tidyverse = ggplot(tomato_data_tidyverse, aes(x=IFE1)) + 
  geom_histogram(binwidth = 1, fill=I("cornflowerblue"), col = I("black")) +
  labs(title = 'Using tidyverse')

hist_pusa_tidyverse = ggplot(tomato_data_tidyverse, aes(x=PusaEarlyDwarf)) + 
  geom_histogram(binwidth = 2, fill=I("gold"), col = I("black")) +
  labs(title = 'Using tidyverse')


ggarrange(hist_ife_base, hist_pusa_base, 
          hist_ife_tidyverse, hist_pusa_tidyverse, ncol = 2, nrow = 2)

```
\pagebreak


## Appendix
```{r Appendix, ref.label=c("setup","sensory_import", "sensory_base", "summary_sensory_base", "sensory_tidyverse", "summary_sensory_tidayverse", "sensory_plots1", "sensory_plots2", "sensory_plots3", "medal_import", "medal_base", "summary_medal_base", "medal_tidyverse",  "summary_medal_tidayverse", "medal_plots", "brain_import", "brain_base", "summary_brain_base", "brain_tidyverse", "summary_brain_tidayverse", "brain_plots", "tomato_import", "tomato_base", "summary_tomato_base", "tomato_tidyverse", "summary_tomato_tidayverse", "tomato_plots"), echo=TRUE, eval=FALSE, tidy=FALSE, include=T}
  
```

