---
title: "HW2_jaeyounglee"
author: "Jaeyoung Lee"
date: \today
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
```
## Problem 3

First of all, it is to handle a mistake. Also, one can handle various versions of a code and see the history of a code. Furthermore, using version control, it is easy to collaborate with others. This is because one can share a code and work on the cloud such as GitHub.



## Problem 4

For each dataset, you should perform the cleaning 2x: first with base R functions (ie no dplyr, piping, etc), second using tidyverse function.  Make sure you weave your code and text into a complete description of the process and end by creating a tidy dataset describing the variables, create a summary table of the data (summary, NOT full listing), note issues with the data, and include an informative plot.  


a. Sensory data from five operators.
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat> 
```{r sensory_import}
######## Sensory data ########
# Getting "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
url_sensory <- "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensory_rawdata <- fread(url_sensory, fill = TRUE, skip = 2, data.table = FALSE)
saveRDS(sensory_rawdata, 'sensory_rawdata.RDS')
sensory_rawdata <- readRDS('sensory_rawdata.RDS')

```

There are missing values in the raw data and the categories "Items" are in the data like oberservations. We need to remove missing values and extract the 'Item' numbers from the data.

```{r sensory_base}
# Using base R function only
# Convert data.frame to matrix and transpose the raw data
matrix_sensory <- t(as.matrix(sensory_rawdata)) 

# Find where the missing values are
na <- which(is.na(matrix_sensory==TRUE))        

# The indexes where Item numbers are in the data
x <- 1
item <- x
for (i in 1:9){
  x <- x+18
  item <- c(item, x)
}

# Remove missing values and 'Item's from the data
sensory_data <- t(matrix(matrix_sensory[-c(na,item)], byrow = T, nrow = 10))  
sensory_data <- data.table(sensory_data)                                      
colnames(sensory_data) <- paste('Item', 1:10) # Assign column names
Opr <- rep(paste('Opr', 1:5), 3)              # Operator names
sensory_data <- cbind(Opr,sensory_data)       # Bind Operator names and the data

# Re-order the rows by names of operators
sensory_data <- sensory_data[order(sensory_data$Opr)]                         
sensory_data_base <- sensory_data

# Final tidy data with base R functions
sensory_data_base

```

Above is the converted tidy data frames using the base R functions only. A summary of the data is as follows:

```{r summary_sensory_base, echo = FALSE}
knitr::kable(summary(sensory_data_base))
```

Now, handle the same data with `tidyverse` package.

```{r sensory_tidyverse}
# Sensory data with tidyverse package
# Making matrix which is the same with base R function but using pipes.
matrix_sensory <- sensory_rawdata %>% as.matrix() %>% t() 
na <- which(is.na(matrix_sensory==TRUE))     # Find missing values

# The indexes where Item numbers are in the data
x <- 1
item <- x
for (i in 1:9){
  x <- x+18
  item <- c(item, x)
}

# Remove missing values and Item numbers from the data
sensory_data <- matrix_sensory[-c(na,item)] %>% matrix(byrow = T, nrow = 10) %>% t()
sensory_data <- data.table(sensory_data)
Opr <- rep(paste('Opr', 1:5), 3)
sensory_data <- bind_cols(Opr,sensory_data) # bind operators and data
colnames(sensory_data) <- c('Opr',paste('Item', 1:10))

# Re-order the rows by names of operators
sensory_data <- sensory_data[order(sensory_data$Opr)] 
sensory_data_tidyverse <- sensory_data

# Final tidy data with tidyverse
sensory_data_tidyverse

```

The result by tidyverse is the same with the base R function. The summary of the data converted by tidyverse is as follows.

```{r summary_sensory_tidayverse, echo = FALSE}
knitr::kable(summary(sensory_data_tidyverse))
```



b. Gold Medal performance for Olympic Men's Long Jump, year is coded as 1900=0.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat>  
```{r medal_import}
######## Long Jump data ########
# Getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
url_medal <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
medal_rawdata <- fread(url_medal)
saveRDS(medal_rawdata, 'medal_rawdata.RDS')
medal_rawdata <- readRDS('medal_rawdata.RDS')

```

The raw data has missing values and wide type data. It is better to reshape the data.
Also, we need two vectors : 'Year' and 'Long Jump'.

```{r medal_base}
# Using base R function only
# Year is coded as 1900 = 0
# Extract Year and Long Jump vectors
year <- c(medal_rawdata[[1]], medal_rawdata[[3]], 
          medal_rawdata[[5]], medal_rawdata[[7]]) + 1900
longjump <- c(medal_rawdata[[2]], medal_rawdata[[4]], 
              medal_rawdata[[6]], medal_rawdata[[8]])

# Bind the vectors as a data table and rename the categories
medal_data <- data.table(year[1:(length(year)-2)], 
                         longjump[1:(length(longjump)-2)])
colnames(medal_data) <- c('Year', 'Long Jump')
medal_data_base <- medal_data

# Final tidy data with base R functions
medal_data_base

```

Above is the converted tidy data frames using the base R functions. A summary of the data is as follows:

```{r summary_medal_base, echo = FALSE}
knitr::kable(summary(medal_data_base))
```

Now, handle the same data with `tidyverse` package.

```{r medal_tidyverse}
# Using tidyverse package
# Year is coded as 1900 = 0
medal_data <- medal_rawdata[,1:8] # remove missing values only columns

# Extracting 'Year' columns and 'Long Jump' columns
colnames(medal_data) <- paste(rep(c('Year', 'Jump'),4), rep(1:4,each = 2))
year <- medal_data[,c(1,3,5,7)] %>% 
  gather(key = 'name1', value = 'Year', 1,2,3,4) %>% filter(Year != na)
year[,2] <- year[,2] + 1900
jump <- medal_data[,c(2,4,6,8)] %>% 
  gather(key = 'name2', value = 'LongJump', 1,2,3,4) %>% filter(LongJump != na)

# Bind the vectors as a data table and rename the categories
medal_data <- bind_cols(year[,2], jump[,2])
colnames(medal_data) <- c('Year', 'Long Jump')
medal_data_tidyverse <- medal_data

# Final tidy data with tidyverse
medal_data_tidyverse


```

The result by tidyverse is the same with the base R function. The summary of the data converted by tidyverse is as follows.

```{r summary_medal_tidyverse, echo = FALSE}
knitr::kable(summary(medal_data_tidyverse))
```







c. Brain weight (g) and body weight (kg) for 62 species.    
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat>  
```{r brain_import}
######## Brain weight data ########
# Getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
url_brain <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
brain_rawdata <- fread(url_brain)
saveRDS(brain_rawdata, 'brain_rawdata.RDS')
brain_rawdata <- readRDS('brain_rawdata.RDS')
```

The data needs two columns which are 'Body Wt' and 'Brain Wt'.

```{r brain_base}
# Using base R function only
# The method is the same with the data from part (b)
# Extract Body Wt and Brain Wt vectors
bodywt <- c(brain_rawdata[[1]], brain_rawdata[[3]], brain_rawdata[[5]])
brainwt <- c(brain_rawdata[[2]], brain_rawdata[[4]], brain_rawdata[[6]])

# Remove missing values
brain_data <- data.table(bodywt[-length(bodywt)], brainwt[-length(brainwt)])
colnames(brain_data) <- c('Body Wt', 'Brain Wt')
brain_data_base <- brain_data

# Final tidy data with base R functions
brain_data_base

```

Above is the converted tidy data frames using the base R functions. A summary of the data is as follows:

```{r summary_brain_base, echo = FALSE}
knitr::kable(summary(brain_data_base))
```

Now, handle the same data with `tidyverse` package.

```{r brain_tidyverse}
# Tidy data with tidyverse
# Remove vectors which have only missing values
brain_data <- brain_rawdata[,1:6]

# Extracting 'Year' columns and 'Long Jump' columns
colnames(brain_data) <- paste(rep(c('bw', 'brw'),3), rep(1:3,each = 2))
bw <- brain_data[,c(1,3,5)] %>% gather(key = 'name1', value = 'BW', 1,2,3) 
brw <- brain_data[,c(2,4,6)] %>% gather(key = 'name2', value = 'BRW', 1,2,3) 

# Bind the vectors as a data table and rename the categories
brain_data <- bind_cols(bw[,2], brw[,2])
colnames(brain_data) <- c('Body Wt', 'Brain Wt')
brain_data_tidyverse <- brain_data

# Final tidy data with tidyverse
brain_data_tidyverse

```

The result by tidyverse is the same with the base R function. The summary of the data converted by tidyverse is as follows.

```{r summary_brain_tidyverse, echo = FALSE}
knitr::kable(summary(brain_data_tidyverse))
```


d. Triplicate measurements of tomato yield for two varieties of tomatos at three planting densities.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat>  
```{r tomato_import} 
######## Tomato data ########
# Getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
url_tomato <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
tomato_rawdata <- fread(url_tomato, skip = 1)
saveRDS(tomato_rawdata, 'tomato_rawdata.RDS')
tomato_rawdata <- readRDS('tomato_rawdata.RDS')

```

The values are grouped in the cells of the data above. Therefore, we should split the cells into single values. 

```{r tomato_base}
# Using base R function only
# Need to split the values
cells <- strsplit(unlist(tomato_rawdata), split = ',', fixed = T) # split the data
categories <- unlist(c(cells[1],cells[2]))  # two categories
values <- as.numeric(unlist(c(cells[3:8]))) # numerical data

# Combine the split values into data frame
tomato_matrix <- matrix(values, byrow = T, ncol = 3)
tomato_matrix <- t(cbind(tomato_matrix[1:2,], tomato_matrix[3:4,], tomato_matrix[5:6,]))

# Bind the data with the densities (categories)
tomato_data <- data.frame(tomato_matrix, as.character(rep(c(10000,20000,30000), each=3)))
colnames(tomato_data) <- c(categories, 'Density')
tomato_data_base <- tomato_data

# Final tidy data with base R functions
tomato_data_base

```
Above is the converted tidy data frames using the base R functions. A summary of the data is as follows:

```{r summary_tomato_base, echo = FALSE}
knitr::kable(summary(tomato_data_base))
```

Now, handle the same data with `tidyverse` package.

```{r tomato_tidyverse}
# Using tidyverse package
# Need to split the values
tomato_data <- tomato_rawdata[,-1] %>% 
  separate(col = '10000', into = c("1","2","3"), sep = ",", convert = T) %>%
  separate(col = '20000', into = c("4","5","6"), sep = ",", convert = T) %>%
  separate(col = '30000', into = c("7","8","9"), sep = ",", convert = T) %>% 
  as.matrix() %>% t()
dens <- rep(c(10000,20000,30000), each = 3) %>% as.character() # Densities

# Bind the data with the densities vector
tomato_data <- tomato_data %>% cbind(dens) %>% as.data.table()
colnames(tomato_data) <- tomato_rawdata[,1] %>% unlist() %>% c("Density")
tomato_data_tidyverse <- tomato_data

# Final tidy data with tidyverse
tomato_data_tidyverse

```


```{r summary_tomato_tidyverse, echo = FALSE}
knitr::kable(summary(tomato_data_tidyverse))
```



## Problem 5

Finish this homework by pushing your changes to your repo.  In general, your workflow for this should be:  

1. git pull -- to make sure you have the most recent repo  
2. In R: do some work  
3. git add -- this tells git to track new files  
4. git commit -- make message INFORMATIVE and USEFUL  
5. git push -- this pushes your local changes to the repo  

If you have difficulty with steps 1-5, git is not correctly or completely setup.  See me for help.

**Only submit the .Rmd and .pdf solution files.  Names should be formatted HW2_lastname.Rmd and HW2_lastname.pdf**

## Optional preperation for next class:  

TBD


## Appendix
```{r Appendix, ref.label=c("setup","sensory_import", "sensory_base", "summary_sensory_base", "sensory_tidyverse", "summary_sensory_tidayverse", "medal_import", "medal_base", "summary_medal_base", "medal_tidyverse", "summary_medal_tidayverse", "brain_import", "brain_base", "summary_brain_base", "brain_tidyverse", "summary_brain_tidayverse", "tomato_import", "tomato_base", "summary_tomato_base", "tomato_tidyverse", "summary_tomato_tidayverse"), echo=TRUE, eval=FALSE, tidy=FALSE, include=T}
  
```

